---
title: "Review-1"
author: "Jameson Watts, Ph.D."
date: "10/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Topics

1. Basic R scripting

Let's say, I wanted to know what 5 + 4 = `r 5+4`

2. Markdown

#### Here's a heading

- create and **unordered** list
- that says *stuff*

3. Philosopy of Tidyverse
4. Filtering

```{r}
library(tidyverse)
wine <- read_csv("../resources/winemag-data.csv")
```


```{r}
wine %>% 
  filter(variety=="Chardonnay") %>% 
  glimpse()
```

5. Arranging
6. Selecting
```{r}
wine %>% 
  arrange(desc(points), price) %>% 
  select(points, price, title)
```

7. Logical Operators
8. Top_n

```{r}
wine %>% 
  top_frac(0.001, price) %>% 
  arrange(desc(price))
```

9. Saving dataframes

```{r}
wine2 <- wine %>% 
  filter(province=="Oregon")
 
wine2
```


10. Pivot wide and long

11. Group_by, counting and summarizing


12. Joining


13. Mutation
14. Strings
15. if_else() and case_when()
16. Visualization basics and plot types