---
title: "Advanced Visualization"
author: "Jameson Watts, Ph.D."
output:
  ioslides_presentation:
    smaller: yes
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
setwd("~/GDrive/teaching/GSMDS-5001/")
options("kableExtra.html.bsTable" = T)
options(knitr.table.format = "html") 
library(tidyverse)
library(moderndive)
library(kableExtra)
d <- read_csv("resources/ab-data.csv")
```

## Agenda

1. Modeling Review (and upgrade)
2. Facets 
3. Random cool stuff

## But first...

### Final project check-in

- **11/2:** You have a data set and have started to look at variables and some descriptive statistics
- **11/16:** You have a comprehensive set of tables and graphs that adequately describe your data and tell a "model-free" story
- **11/30:** You have finished modeling the relationships in the data and are confident in the qualitiy and validity of the results
- **12/14:** All writing is done, all tables and graphs are clear and sized correctly, report knits to a beautiful html presentation suitable for consumption by a non-technical manager

# Regression review 

## Topics covered

1. Univariate linear regression
2. Hypothesis testing for proportions and continuous outcomes
3. Multiple regression
4. **Univariate regression as alternative to t-test**
5. **Categorical x categorial interactions**
7. **Multiple regresson for uplift modeling**
8. **Regression diagnostics**
9. **Categorical x continuous interactions**

## Practice: Univariate Regression for A/B

#### What is the difference in open and click rates for email A vs. email B? 
\
\
Use linear regression (not proportionality tests) to answer the question.

## Solution
```{r}
rd <- d %>% 
  filter(group!="ctrl") %>% 
  mutate(email_A=(group=="email_A"))
```

Opens
```{r}
m <- lm(open~email_A, data = rd)
summary(m)$coef
```

Clicks
```{r}
m <- lm(click~email_A, data = rd)
summary(m)$coef

```


## Compared to proportionality test?
```{r}
success <- d %>% 
  filter(group!="ctrl") %>% 
  group_by(group,click) %>% 
  count() %>% 
  filter(click==1)
trials <- d %>% 
  filter(group!="ctrl") %>% 
  count(group)

prop.test(success$n,trials$n)
```

## Practice: categorical interactions

#### What is the difference in the impact that our promotional email had on purchases for those with past purchases greater than zero?
\
\
Use linear regression with a categorical by categorical interaction.

## Solution
```{r}
rd <- d %>% 
  mutate(email=(group!="ctrl"))

m <- lm(purch~email*(past_purch>0), data = rd)
summary(m)$coef
```

## Compare that with filter and t-test
```{r}
t.test(purch~email, data=filter(rd, past_purch>10))
```

## Visualize it (upgrade)

```{r, fig.height=3}
rd %>% 
  mutate(pp=(past_purch>0)) %>% 
  group_by(email, pp) %>% 
  summarise(purch = mean(purch)) %>% 
  ggplot() +
  aes(x = pp, y = purch, color = email) +
  geom_line(aes(group = email)) +
  geom_point()
```


## Practice: uplift modeling

#### Given the model used in the previous practice problem, how does recieving an email change the predicted purchase amount for someone with the same baseline variables as customer 1001?

## Solution

```{r}
m <- lm(purch~email*(past_purch>0), data = mutate(d,email=(group!="ctrl")))
(pred <- predict(m, cbind(email=c(TRUE, FALSE), slice(d,1001))))
(lift <- pred[1] - pred[2])
```

The model predicts a lift of \$`r round(lift,2)` for customer 200.

## Practice: regression diagnostics

#### Should we use a model with (past_purchase>0) as the cutoff or simply past_purch (as a continuous variable)? Split the data into test and train subsets and evaluate the models on their predictive quality.

## Solution - models and subsets

```{r}
# Randomly shuffle order of rows:
d_shuffled <- d %>%
  mutate(email=(group!="ctrl")) %>% 
  sample_frac(size = 1, replace = FALSE)
# Split into train and test:
split <-  round(.8*nrow(d)) # 80% of dataframe
train <- d_shuffled %>%
  slice(1:split)
test <- d_shuffled %>%
  slice(split+1:nrow(d))

m1 <- lm(purch~email*(past_purch>0), data = train)
m2 <- lm(purch~email*past_purch, data = train)
```

## Solution - rmse

Model 1: past_purch > 0
```{r}
get_regression_points(m1, newdata = test) %>% 
  mutate(sq_residuals = residual^2) %>%
  summarize(rmse = sqrt(mean(sq_residuals)))
```
Model 2: past_purch continous
```{r}
get_regression_points(m2, newdata = test) %>% 
  mutate(sq_residuals = residual^2) %>%
  summarize(rmse = sqrt(mean(sq_residuals)))
```

## Practice: categorical by continuous interactions

```{r}
m <- lm(purch~email*past_purch, data = train)
summary(m)$coef
```

What effect does email have on the relationship between past purchases and current purchases?

## Visualising the interaction (upgrade)

```{r, fig.height=4}
train %>% 
  ggplot(aes(past_purch, purch, color=email))+
  geom_smooth(method=lm)+
  xlim(0,500)
```


## Continuous by continuous interactions (upgrade)

```{r}
m <- lm(purch~last_purch*visits, data = train)
summary(m)$coef
```

## Visualizing the interaction (upgrade)

```{r include=FALSE}
library(margins)
sample <- sample_frac(d, size = .03, replace = FALSE)
m <- lm(purch~last_purch*visits, data = sample)

# Create a range of meaningful values of x for the prediction. We don't need all of them, 
#  so I am going to increment by .5
x.seq <- seq(from = round(min(sample$last_purch)), to = round(max(sample$last_purch)), by = 5)

# Now we get the low and high values of m based on one standard deviation 
#  above and below the mean
low.v <- mean(sample$visits) - sd(sample$visits)
high.v <- mean(sample$visits) + sd(sample$visits)

# Lastly we create two new dataframes based on our new values of x and our low/high values of m
low.df <- data_frame(last_purch = x.seq, visits = low.v)
high.df <- data_frame(last_purch = x.seq, visits = high.v)

predict.low <- cplot(m, x = "last_purch", what = c("prediction"), data = low.df, draw = FALSE) %>% 
  mutate(visits=-1)
predict.high <- cplot(m, x = "last_purch", what = c("prediction"), data = high.df, draw = FALSE) %>% 
  mutate(visits=1)
predict.model <- bind_rows(predict.low, predict.high) %>%
  mutate(visits = factor(visits, levels = c(-1, 1), labels = c("Low Visits", "High Visits")))

predict.plot <- ggplot(data = predict.model, aes(x = xvals, group = visits)) + 
  geom_line(aes(y = yvals, color = visits)) + 
  geom_ribbon(alpha = .2, aes(ymin = lower, ymax = upper)) + 
  labs(title = "Effect of last_purch on purch Moderated by visits",
       subtitle = "Simple Slopes at +/- 1 Standard Deviation of Mean-Centered Visits",
       y = "Predicted Values of Purch",
       x = "Last Purch")+
      theme(legend.position = c(.7, .85))

mm <- cplot(m, dx="last_purch", x="visits", what="effect", data = sample, draw = F)
margins.plot <- ggplot(mm, aes(x = xvals, y = yvals)) + 
    geom_line(color = "red") + 
    geom_ribbon(alpha = .2, aes(ymin = lower, ymax = upper)) + 
    geom_hline(yintercept = 0, linetype = "dashed") +  # I add a vertical line at zero to help with interpretation
    labs(title = "Marginal Effect of last_purch on purch as a function of visits",
         subtitle = "Marginal Effect of last_purch Across Range of Mean-Centered visits",
         y = "Estimated Effect of last_purch on purch",
         x = "Visits")
```


```{r, echo=F}
library(cowplot)
plot_grid(predict.plot, margins.plot, ncol = 2, align = "h")
```

# Faceting

## Setup
```{r}
wine <- read_csv("../resources/winemag-data.csv") %>% 
  filter(!is.na(price)) %>%
  mutate(year = as.numeric(str_extract(title,"(\\d{4})"))) %>% 
  filter(province %in% c("Washington","Oregon","California")) %>% 
  filter(variety %in% c("Pinot Noir", "Chardonnay", "Cabernet Sauvignon"))
```

## Facet wrap

```{r}
wine %>% 
  ggplot(aes(points, fill=province))+
  geom_density()+
  facet_wrap(facets = ~province)
```


## Facet grid

```{r}
wine %>% 
  ggplot(aes(points, fill=province))+
  geom_density()+
  facet_grid(. ~ province)
```

## Facet grid horizontal

```{r}
wine %>% 
  ggplot(aes(points, fill=province))+
  geom_density()+
  facet_grid(province ~ .)
```

## Facet grid multiple

```{r}
wine %>% 
  ggplot(aes(points, fill=variety))+
  geom_density()+
  facet_grid(province~variety)
```

## Exercise

1. Create facet grid of transparent jittered points 
2. by province and variety
3. that shows log(price) as a function of points
4. with a regression line

## Solution

```{r}
wine %>% 
  ggplot(aes(points,log(price), color=variety))+
  geom_jitter(alpha=0.2)+
  geom_smooth(method = "lm", color="black")+
  facet_grid(province~variety)
```

## Cowplot setup

```{r}
library(cowplot)
plot1 <- ggplot(wine, aes(points,log(price), color=variety))+
  geom_jitter(alpha=0.2)+
  geom_smooth(method = "lm", color="black")+
  facet_grid(province~variety)+
  theme(legend.position="none")
plot2 <- ggplot(wine, aes(points, fill=variety))+
  geom_density()+
  facet_grid(province~variety)+
  theme(legend.position="none")
```

## Cowplot plot
```{r}
plot_grid(plot1, plot2, ncol = 2, align = "h")
```

# Random cool stuff

## Custom themes

- element_text()
- element_rect()
- element_line()
- element_blank()

```{r}
library(wesanderson)
jameson <- function(){
  list(
    theme_minimal() +
    theme(
      text = element_text(family = "Palatino", color = "gray25"),
      plot.subtitle = element_text(size = 12),
      plot.caption = element_text(color = "gray30"),
      plot.background = element_rect(fill = "gray95"),
      plot.margin = unit(c(5, 10, 5, 10), units = "mm")
    ),
    scale_color_manual(values = wes_palette(n=5, name="Moonrise3", type = "continuous")),
    scale_fill_manual(values = wes_palette(n=5, name="Moonrise3", type = "continuous"))
  )
}

```

## Using the 'jameson' theme

```{r}
wine %>% 
  ggplot(aes(points, fill=variety))+
  geom_density()+
  facet_grid(province~variety)+
  jameson()
```


## Dot plot

```{r}
wine %>% 
  filter(year %in% c(2000, 2010)) %>% 
  group_by(province,year) %>% 
  summarise(points=mean(points)) %>% 
  ggplot(aes(x = points, y = province))+
  geom_path(arrow = arrow(length = unit(1.5, "mm"), type = "closed"))+
  labs(title = "Change in average points from 2000 to 2010 for west coast states") +
  jameson()
  
```


## Overplotting

```{r}
wine %>% 
  ggplot(aes(points, log(price), color=variety))+
  geom_jitter()+
  facet_grid(. ~ variety)+
  jameson()+
  theme(legend.position="none")
```


## 2d density

```{r}
wine %>% 
  ggplot(aes(points, log(price), color=variety))+
  geom_density_2d()+
  facet_grid(. ~ variety)+
  jameson()+
  theme(legend.position="none")
```

## 2d Histogram

```{r}
wine %>% 
  ggplot(aes(points,log(price)))+
  geom_bin2d(bins=20)+
  facet_grid(. ~ variety)+
  theme_minimal() +
    theme(
      text = element_text(family = "Palatino", color = "gray25"),
      plot.subtitle = element_text(size = 12),
      plot.caption = element_text(color = "gray30"),
      plot.background = element_rect(fill = "gray95"),
      plot.margin = unit(c(5, 10, 5, 10), units = "mm")
    )
```

## Long exercise

1. Gather in groups of 2-3
2. Create the ugliest custom theme that you can
3. Apply it to a plot of the wine data (you can use full dataset)
4. To show something **completely new!!!**


