---
title: "Foundations of Data Science with R"
author: "Jameson Watts, Ph.D."
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Agenda

1. Course Overview and Expectations
2. Example Analysis of Wine Prices

# Course Overview and Expectations

## Contact Information
- Email: jwatts\@willamette.edu
- Office Hours: after class 
- Appointments: jamesonwatts.youcanbook.me (Skype or phone call)

## Class Materials

- [Base R](https://cran.r-project.org/)
- [RStudio 1.2](https://www.rstudio.com/products/rstudio/download)
- [R for Data Science](https://r4ds.had.co.nz)
- [DataCamp Classroom](https://www.datacamp.com/groups/shared_links/526ee016c61b2912b2bbc7a2f7fd1f1cf1ea3d09)

Other resources:

- https://twitter.com/r4dscommunity
- https://bookdown.org/yihui/rmarkdown
- http://google.com

## Reading the Course Outline

- Class Topics
  + Subjects I plan to cover during that day's lecture
- Reading and Assignments
  + DCC: assignments in the DataCamp Classroom
  + R4ds: chapters to read in the online textbook
  
## Assignments
- DataCamp homework assignments (25%)
- Midterm exams (50%)
- Final Presentations and Report (25%)

## Course Policies and Expectations
- Name tents
- Collaboration
- Late work
- Effort
  + 2-4 hours outside of class each week
  + struggle, Google, struggle, doh!
  + start with the basics... ramp up very fast

# Analysis of wine prices

## Dataset



```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

