---
title: "Summarize and Plot"
author: "Jameson Watts, Ph.D."
date: "10/3/2019"
# output: html_document
output:
  ioslides_presentation:
    smaller: yes
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
library(tidyverse)
```

## Agenda

1. Expectations for the mid-term
2. Review (and upgrade) 
3. Visualization techniques

# Mid-term

## Expectation and format

1. 45 minutes of review
2. 120 minutes to complete
3. Must be submitted as both Rmd file and knitted HTML
4. Open everything (notes, book, internet)... except communication
5. Questions will get progressively more difficult

# Review

## Joining, Mutation, and Strings

First things first, let's load in my new data on monthly rainfall and take a look. What do you notice?
```{r, warning=FALSE, message=FALSE}
rain <- read_csv("../resources/rainfall.csv")
rain
```

## Tidyr 1.0.0

Tidyr has replaced spread() and gather() with pivot_wider() and pivot_longer(). I encourage you to read about the developments [here](https://www.tidyverse.org/articles/2019/09/tidyr-1-0-0/).

```{r}
rain %>% 
  rename("year"="Year") %>% 
  pivot_longer(-year,names_to = "month", values_to = "rainfall")
```

## Exercise

1. Load in the wine data
2. Get rid of prices that are NA
3. Only keep Oregon wines
4. Extract the year from the title (as numeric)
5. Join with rainfall data
6. Pivot longer

## Solution

*Note:* See [here](https://github.com/batpigandme/tidyexplain/tree/pivot) for an intuitive tutorial on joins.

```{r}
wine <- read_csv("../resources/winemag-data.csv") %>% 
  filter(!is.na(price)) %>%
  filter(province=="Oregon") %>% 
  mutate(year = as.numeric(str_extract(title,"(\\d{4})"))) %>% 
  left_join(rain, by=c("year"="Year")) %>% 
  pivot_longer(16:27,names_to = "month", values_to = "rainfall")

wine %>% 
  select(title, month, year, rainfall)
```

## Skills upgrade: all about case_when()

Sometimes you want to do a bunch of if/else in your mutate all at once.

```{r, eval=FALSE}
wine %>% 
  mutate(month_number =
    case_when(
      month=="Jan" ~ 1,
      month=="Feb" ~ 2,
      month=="Mar" ~ 3,
      month=="Apr" ~ 4,
      month=="May" ~ 5,
      month=="Jun" ~ 6,
      month=="Jul" ~ 7,
      month=="Aug" ~ 8,
      month=="Sep" ~ 9,
      month=="Oct" ~ 10,
      month=="Nov" ~ 11,
      month=="Dec" ~ 12,
    )
  )
```

## Exercise

1. Partner up and choose a driver 
2. Use case_when() and/or str_detect() with regular expressions
3. To create a new variable called "character"
4. With values of 'tart,' 'spicy,' 'bold' and 'cherry' 
5. For Oregon wines with those terms in their description
6. Then plot the density of log(price) by character

*Hint:* you may want to pivot_wider() first

## Solution (code)

```{r}
wine <- wine %>%
  pivot_wider(names_from = month, values_from = rainfall) %>% 
  mutate(character=
    case_when(
      str_detect(description,"[Tt]art") ~ 'tart',
      str_detect(description,"[Sp]icy") ~ 'spicy',
      str_detect(description,"[Bb]old") ~ 'bold',
      str_detect(description,"[Cc]herry") ~ 'cherry'
    )
  ) 
```

## Solution (graph)

```{r}
wine %>% 
  filter(!is.na(character)) %>% 
  ggplot(aes(log(price), fill=character))+
    geom_density(alpha=.7)
```



# Visualization basics

## Overview

- Aesthetics
  + x = 
  + y =
  + fill = 
  + color =
- Geometry
  + Line plots
  + Bar plots
  + Histograms
  + Violin plots

## Types of plots

- Line and scatter
  + geom_point()
  + geom_jitter()
  + geom_line()
- Bar
  + geom_bar()
  + geom_col()
- Histograms
  + geom_histogram()
  + geom_density()
- Box
  + geom_box()
  + geom_voilin()


## Scatter plot example
```{r}
wine %>% 
  filter(year>1995) %>% 
  filter(!is.na(character)) %>% 
  ggplot(aes(x=year, y=points, color=character)) +
    geom_jitter()
```

## Bar plot example
```{r}
wine %>% 
  filter(!is.na(character)) %>% 
  ggplot(aes(character))+
    geom_bar()
```

## Exercise

Plot the counts of each character of wine between 1995 and 2015

## Solution

```{r, fig.height=3}
wine %>% 
  filter(year>1995 & year <= 2015) %>%
  filter(!is.na(character)) %>% 
  group_by(year,character) %>% 
  summarise(char_cnt=n()) %>% 
  ggplot(aes(year,char_cnt, color=character))+
    geom_line()
```

## Histogram example

```{r}
wine %>% 
  ggplot(aes(price))+
    geom_histogram()
```

## Violin Plots
```{r}
wine %>% 
  filter(!is.na(character)) %>% 
  ggplot(aes(character,log(price), color=character))+
    geom_violin()
```


## Long Exercise

Use any of the techniques that you've learned thus far to answer the following:

*Is there a relationship between rainfall and wine quality in Oregon?*

## One simple solution (and some bonus code)

```{r}
rains <- rain %>%
  rename("year"="Year") %>% 
  pivot_longer(-year,names_to = 'month',values_to = 'rainfall') %>% 
  mutate(rainfall=ifelse(is.na(rainfall),0,rainfall)) %>%
  filter(month %in% c('May','Jun','Jul','Aug','Sep')) %>% #note the %in% operator
  group_by(year) %>% 
  summarise(summer_rain=sum(rainfall))

wines <- wine %>% 
  filter(points > 88) %>% 
  group_by(year) %>% 
  summarize(avg_price=mean(price), avg_points=mean(points)) %>% 
  left_join(rains)
```

## And a graph


```{r}
wines %>% 
  ggplot(aes(log(summer_rain),avg_points))+
    geom_point()+
    geom_smooth(method = lm)
```



