---
title: "What Makes Oregon Wine Unique?"
author: "Jameson Watts, Ph.D."
date: "10/30/2019"
output:
  ioslides_presentation:
    widescreen: yes
    incremental: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)

library(tidyverse)
library(tidytext)
library(scales)

wine <- read_csv("../resources/winemag-data.csv") %>% 
  filter(!is.na(price)) %>%
  mutate(year = as.numeric(str_extract(title,"(\\d{4})")))

data(stop_words)
wtxt <- wine %>% 
  unnest_tokens(word, description) %>% 
  anti_join(stop_words) %>% 
  filter(str_detect(string = word, pattern = "[a-z+]")) %>%  # get rid weird non alphas
  filter(str_length(word)>3) %>%  # get rid of strings shorter than 3 characters
  group_by(variety, word) %>% 
  mutate(total=n()) %>% 
  ungroup()

west_coast <- c("Washington","Oregon","California")
pinots <- c("Burgundy","California","Oregon")
```


## Pinot Noir !!!

### Right??
\
\
But wait ...is there evidence?

## Top 5 wine varieties on the west coast
```{r}
top_5 <- (wine %>%
  filter(province %in% west_coast) %>%
  count(variety) %>% 
  top_n(5,n))$variety 

wine %>% 
  filter(province %in% west_coast) %>%
  filter(variety %in% top_5) %>% 
  ggplot(aes(x=province, fill=variety)) +
    geom_bar()
```

## Wines at 98 points or higher on the west coast
```{r}
wine %>%
  filter(points >= 98) %>%
  filter(province %in% west_coast) %>%
  ggplot(aes(x=province, fill=variety)) +
geom_bar()
```

## Oregon's Top wines 
```{r}
wine %>% 
  filter(province == "Oregon") %>% 
  filter(points >= 97) %>% 
  arrange(desc(points),desc(price)) %>% 
  select(winery,variety,points,price)
```

# So what really makes us unique?

## Different from California Pinot...

```{r}
wtxt %>% 
    filter(province=="Oregon" | province=="California") %>% 
    filter(variety=="Pinot Noir") %>% 
    filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling"))) %>% 
    filter(total > 400) %>% 
    group_by(province, word) %>%
    count() %>% 
    group_by(province) %>% 
    mutate(proportion = n / sum(n)) %>% 
    pivot_wider(id_cols = word, names_from = province, values_from = proportion) %>% 
    ggplot(aes(x = Oregon, y = California, color = abs(Oregon - California))) +
      geom_abline(color = "gray40", lty = 2) +
      geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
      geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
      scale_x_log10(labels = percent_format()) +
      scale_y_log10(labels = percent_format()) +
      scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
      theme(legend.position="none") +
      labs(x = "Oregon", y = "California", title = "Words describing Pinot Noir from California and Oregon")
```

## Different from French Burgundy
```{r, warning=FALSE}
wtxt %>% 
  filter(province=="Oregon" | province=="Burgundy") %>% 
  filter(variety=="Pinot Noir") %>% 
  filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling","bottle","finish"))) %>% 
  filter(total > 400) %>% 
  group_by(province, word) %>%
  count() %>% 
  group_by(province) %>% 
  mutate(proportion = n / sum(n)) %>% 
  pivot_wider(id_cols = word, names_from = province, values_from = proportion) %>% 
  ggplot(aes(x = Oregon, y = Burgundy, color = abs(Oregon - Burgundy))) +
    geom_abline(color = "gray40", lty = 2) +
    geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
    geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
    scale_x_log10(labels = percent_format()) +
    scale_y_log10(labels = percent_format()) +
    scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
    theme(legend.position="none") +
    labs(x = "Oregon", y = "Burgundy", title = "Describing Pinot Noir from Burgundy and Oregon")
```

## Unique Oregon pinot descriptors (compared to California)

```{r}
dtxt <- wtxt %>% 
  filter(province=="Oregon" | province=="California") %>% 
  filter(variety=="Pinot Noir") %>% 
  filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling","bottle","finish"))) %>% 
  filter(total > 400) %>% 
  group_by(province, word) %>%
  count() %>% 
  group_by(province) %>% 
  mutate(proportion = n / sum(n)) %>% 
  pivot_wider(id_cols = word, names_from = province, values_from = proportion) %>% 
  mutate(diff=Oregon-California) 

dtxt %>%
  top_n(25, diff) %>%
  mutate(word = reorder(word, diff)) %>%
  ggplot(aes(word, diff)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()
```

## Rating oregon flavors

```{r}
wine %>%
  filter(variety=="Pinot Noir") %>% 
  mutate(character=
    case_when(
      str_detect(description,"fruit") ~ 'fruit',
      str_detect(description,"chocolate") ~ 'chocolate',
      str_detect(description,"berry") ~ 'berry',
      str_detect(description,"tart") ~ 'tart'
    )
  ) %>% 
  filter(!is.na(character)) %>% 
  ggplot(aes(points, fill=character))+
    geom_density(alpha=.5)
```

## Beyond the highest scores...

```{r}
wine %>% 
  filter(variety=="Pinot Noir") %>% 
  filter(province %in% pinots) %>% 
  ggplot(aes(x=province, y=points))+
  geom_boxplot()
  
```

## More consistent Pinots

```{r}
wine %>% 
  filter(variety=="Pinot Noir") %>% 
  filter(province %in% pinots) %>% 
  ggplot(aes(x=points, fill=province))+
  geom_density(alpha=0.5)
  
```

## Average points per year for Pinot Noir

```{r}
wine %>% 
  filter(province %in% pinots) %>% 
  filter(variety=="Pinot Noir") %>% 
  filter(year >= 2000 & year < 2015) %>% 
  group_by(year, province) %>% 
  summarize(avg_points = mean(points)) %>% 
  ggplot(aes(x=as.integer(year), y=avg_points, color=province)) +
    geom_line()
```

## Points per $$ for Pinot Noir

```{r}
wine %>% 
  filter(variety=="Pinot Noir") %>% 
  filter(province %in% pinots) %>% 
  ggplot(aes(x=points, y=log(price), color=province))+
  geom_point(alpha=0.1,show.legend = FALSE)+
  geom_smooth(method = lm,show.legend = FALSE) +
  facet_wrap(~ province, scales="fixed")
```

## More pricing power than California

Increases in points affect wines from the different regions differently.

```{r}

model = list()
for(p in pinots){
  model <- lm(log(price)~points, filter(wine,province==p, variety=="Pinot Noir"))
  pct <- round((exp(coef(model)["points"]) - 1) * 100,2)
  print(str_c("For ",p,", a 1 point ratings increase leads to a ",pct,"% increase in price."))
}
```

## Generally more for your money

```{r}
wine %>% 
  filter(variety=="Pinot Noir") %>% 
  filter(province %in% pinots) %>% 
  mutate(points_per_dollar= points/price) %>% 
  ggplot(aes(x=province, y=log(points_per_dollar)))+
  geom_boxplot()
  
```


## Summary

Oregon Pinot Noir...

1. Is fruit forward, with hints of chocolate and berries
2. Has lower overall variance in quality
3. Has been rising in quality over time
4. Has more pricing power than California
5. Is the best bang for your buck

## Questions?

Quality Oregon wines under $50
```{r}
wine %>% 
  filter(province == "Oregon") %>% 
  filter(variety != "Riesling") %>% 
  filter(price < 50) %>% 
  select(title, points, price) %>% 
  arrange(desc(points), price) %>% 
  slice(1:10)
```

## Thank you!!

...at least we're better than Texas.

```{r}

wine %>% 
  filter(province %in% c("Oregon","Texas")) %>% 
  ggplot(aes(x=points, fill=province))+
  geom_density(alpha=0.3)
  
```

